A New Hybrid Machine Learning Method for Stellar Parameter Inference

### 2025-03-07 paper/paper2/paper.tex
Fix <

\begin{abstract}
    The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is more powerful than traditional techniques.
    Here we introduce a hybrid ML (HML) method combin{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters \textbf{of} stellar spectra.
    We study $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method \textbf{could be extended} to other \textbf{parameters} such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
    We first use \texttt{blase} \textbf{to} semi-empirical\textbf{ly recast} spectra \textbf{into} sets of \textbf{Voigt profiles}.
    \texttt{blase} is \textbf{run} on 1,314 \textbf{synthetic} spectra from \textbf{a rectilinear subset of} the PHOENIX model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0], $\lambda$: [8038, 12849] \AA)}.
    For \textbf{128,723 detected features}, we map stellar parameters to spectral line parameters using linear interpolation.
    \textbf{This} create\textbf{s} the PHOENIX generator, \textbf{enabling parallelized spectral synthesis}.
    Gaussian Process minimization is used to infer stellar parameters by minimizing \textbf{a} root-mean-square loss \textbf{function}.
    Testing \textbf{210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0]), we find inference errors:} $T_{\mathrm{eff}}$\textbf{: 93 K, $\log(g)$: 0.24, and [Fe/H]: 0.056 for $T_\text{eff}$ < 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26, and [Fe/H]: 0.16 for $T_\text{eff} \ge$ 7000 K}.
    \textbf{We also upload online} an archive of \texttt{blase} \textbf{models} of the PHOENIX subset.
    This proof-of-concept \textbf{study} shows that semi-empirical HML is a viable alternative to traditional approaches \textbf{in spectroscopic inference}.
\end{abstract}

@@

\textbf{For $T_{\mathrm{eff}}$ < 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24}, and [Fe/H] \textbf{0.056} \textbf{(standard deviations of 139 K, 0.40, and 0.15, respectively)}.
\textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26, and [Fe/H] 0.16 (standard deviations of 293 K, 0.32, and 0.23, respectively).
We see that \texttt{blase3D} is more effective and generally more self-consistent when predicting stellar parameters of lower temperature spectra.}

@@

We \textbf{then} introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters.} 
\textbf{Two test passes on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex) yield} average absolute deviations from true values of \textbf{93} K in $T_\mathrm{eff}$, \textbf{0.24 dex} in $\log(g)$, and \textbf{0.056} dex in [Fe/H] \textbf{in low-temperature ($T_\mathrm{eff}$ < 7000) regimes.
In high-temperature regimes ($T_\text{eff} \ge$ 7000), the errors increase to 347 K in $T_\mathrm{eff}$, 0.26 dex in $\log(g)$, and 0.16 dex in [Fe/H]}.

>>>>

\begin{abstract}
    The advent of machine learning (ML) is revolutionary to numerous scientific disciplines, with a growing number of examples in astronomical spectroscopic inference, as ML is more powerful than traditional techniques.
    Here we introduce a hybrid ML (HML) method combin{ing} automatic differentiation, interpolation, and Bayesian optimization to infer stellar parameters \textbf{of} stellar spectra.
    We study $T_{\mathrm{eff}}$, $\log(g)$, and [Fe/H], but this method \textbf{could be extended} to other \textbf{parameters} such as [$\alpha$/Fe] \textbf{(alpha element abundance)}, C/O \textbf{(carbon-oxygen ratio)}, and $f_{sed}$ \textbf{(sedimentation efficiency)}.
    We first use \texttt{blase} \textbf{to} semi-empirical\textbf{ly recast} spectra \textbf{into} sets of \textbf{Voigt profiles}.
    \texttt{blase} is \textbf{run} on 1,314 \textbf{synthetic} spectra from \textbf{a rectilinear subset of} the PHOENIX model grid \textbf{($T_\text{eff}$: [2300, 12000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0], $\lambda$: [8038, 12849] \AA)}.
    For \textbf{128,723 detected features}, we map stellar parameters to spectral line parameters using linear interpolation.
    \textbf{This} create\textbf{s} the PHOENIX generator, \textbf{enabling parallelized spectral synthesis}.
    Gaussian Process minimization is used to infer stellar parameters by minimizing \textbf{a} root-mean-square loss \textbf{function}.
    Testing \textbf{210 noise-free models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6], [Fe/H]: [-0.5, 0]), we find inference errors:} $T_{\mathrm{eff}}$\textbf{: 93 K, $\log(g)$: 0.24, and [Fe/H]: 0.056 for $T_\text{eff} < $ 7000 K, and $T_{\mathrm{eff}}$: 347 K, $\log(g)$: 0.26, and [Fe/H]: 0.16 for $T_\text{eff} \ge$ 7000 K}.
    \textbf{We also upload online} an archive of \texttt{blase} \textbf{models} of the PHOENIX subset.
    This proof-of-concept \textbf{study} shows that semi-empirical HML is a viable alternative to traditional approaches \textbf{in spectroscopic inference}.
\end{abstract}

@@

\textbf{For $T_{\mathrm{eff}} < $ 7000 K, the inferred} $T_{\mathrm{eff}}$ differed from the true \textbf{value} by an average of \textbf{93} K, $\log(g)$ \textbf{0.24}, and [Fe/H] \textbf{0.056} \textbf{(standard deviations of 139 K, 0.40, and 0.15, respectively)}.
\textbf{For $T_{\mathrm{eff}} \ge$ 7000 K, the inferred $T_{\mathrm{eff}}$ differed from the true value by an average of 347 K, $\log(g)$ 0.26, and [Fe/H] 0.16 (standard deviations of 293 K, 0.32, and 0.23, respectively).
We see that \texttt{blase3D} is more effective and generally more self-consistent when predicting stellar parameters of lower temperature spectra.}

@@

We \textbf{then} introduce and test our \textbf{GP minimization} algorithm \textbf{to infer stellar parameters.} 
\textbf{Two test passes on 210 noise-free synthetic models ($T_{\mathrm{eff}}$: [3000, 11000] K, $\log(g)$: [2, 6] dex, [Fe/H]: [-0.5, 0] dex) yield} average absolute deviations from true values of \textbf{93} K in $T_\mathrm{eff}$, \textbf{0.24 dex} in $\log(g)$, and \textbf{0.056} dex in [Fe/H] \textbf{in low-temperature ($T_\mathrm{eff} < $ 7000) regimes.
In high-temperature regimes ($T_\text{eff} \ge$ 7000), the errors increase to 347 K in $T_\mathrm{eff}$, 0.26 dex in $\log(g)$, and 0.16 dex in [Fe/H]}.
